---
title: "CaseStudy10_LeeMessaGarza"
authors: "Brian Less, Araya Messa, Tony Garza"
date: "July 24, 2016"
output: 
  html_document:
    keep_md: TRUE
---
# Import and Prepare Data

```{r}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(fpp)



#################################################################################
#   1.1 Import the exports data and the indicators                              #
#################################################################################

### !!! In order to test the script, it is necessary to change the three
### file paths. The files have been sent together with the script!!!

# The Export data for Chulwalar   are in two .csv files.
# One file for the as is data: ImportedAsIsDataChulwalar.csv
# and another one for the plan data: ImportedPlanDataChulwalar.csv

setwd("~/Dropbox/SMUDataScienceClass/1st class/Week 10/CaseStudy10_LeeMessaGarza")

ImportedAsIsData <- read.csv("ImportedAsIsDataChulwalar.csv", header = F, sep=";", fill = T) #chose ImportedAsIsDataChulwalar.csv

ImportedPlanData <- read.csv("ImportedPlanDataChulwalar.csv", header = F, sep=";", fill = T) #chose ImportedPlanDataChulwalar.csv



# The indicators data is also in a file: ImportedIndicatorsChulwalar.csv
ImportedIndicators <- read.csv("ImportedIndicatorsChulwalar.csv", header = F, sep=";", fill = T) # chose ImportedIndicatorsChulwalar.csv


TotalAsIsVector <- c(ImportedAsIsData [2:13,2],ImportedAsIsData [2:13,3],ImportedAsIsData [2:13,4],ImportedAsIsData [2:13,5],ImportedAsIsData [2:13,6],ImportedAsIsData [2:13,7])
EfakAsIsVector <- c(ImportedAsIsData [16:27,2],ImportedAsIsData [16:27,3],ImportedAsIsData [16:27,4],ImportedAsIsData [16:27,5],ImportedAsIsData [16:27,6],ImportedAsIsData [16:27,7])
WugeAsIsVector <- c(ImportedAsIsData [30:41,2],ImportedAsIsData [30:41,3],ImportedAsIsData [30:41,4],ImportedAsIsData [30:41,5],ImportedAsIsData [30:41,6],ImportedAsIsData [30:41,7])
TotalEtelAsIsVector <- c(ImportedAsIsData [44:55,2],ImportedAsIsData [44:55,3],ImportedAsIsData [44:55,4],ImportedAsIsData [44:55,5],ImportedAsIsData [44:55,6],ImportedAsIsData [44:55,7])
BlueEtelAsIsVector <- c(ImportedAsIsData [58:69,2],ImportedAsIsData [58:69,3],ImportedAsIsData [58:69,4],ImportedAsIsData [58:69,5],ImportedAsIsData [58:69,6],ImportedAsIsData [58:69,7])
RedEtelAsIsVector <- c(ImportedAsIsData [72:83,2],ImportedAsIsData [72:83,3],ImportedAsIsData [72:83,4],ImportedAsIsData [72:83,5],ImportedAsIsData [72:83,6],ImportedAsIsData [72:83,7])
YearAsIsVector <- c(ImportedAsIsData [86,2],ImportedAsIsData [86,3],ImportedAsIsData [86,4],ImportedAsIsData [86,5],ImportedAsIsData [86,6],ImportedAsIsData [86,7])
TotalAsIsVector_2014 <- c(ImportedAsIsData[2:13,8])

PlanVector <- c(ImportedPlanData[2:13,2],ImportedPlanData[2:13,3],ImportedPlanData[2:13,4],ImportedPlanData[2:13,5],ImportedPlanData[2:13,6],ImportedPlanData[2:13,7])
EfakPlanVector <- c(ImportedPlanData[16:27,2],ImportedPlanData[16:27,3],ImportedPlanData[16:27,4],ImportedPlanData[16:27,5],ImportedPlanData[16:27,6],ImportedPlanData[16:27,7])
WugePlanVector <- c(ImportedPlanData[30:41,2],ImportedPlanData[30:41,3],ImportedPlanData[30:41,4],ImportedPlanData[30:41,5],ImportedPlanData[30:41,6],ImportedPlanData[30:41,7])
TotalEtelPlanVector <- c(ImportedPlanData[44:55,2],ImportedPlanData[44:55,3],ImportedPlanData[44:55,4],ImportedPlanData[44:55,5],ImportedPlanData[44:55,6],ImportedPlanData[44:55,7])
BlueEtelPlanVector <- c(ImportedPlanData[58:69,2],ImportedPlanData[58:69,3],ImportedPlanData[58:69,4],ImportedPlanData[58:69,5],ImportedPlanData[58:69,6],ImportedPlanData[58:69,7])
RedEtelPlanVector <- c(ImportedPlanData[72:83,2],ImportedPlanData[72:83,3],ImportedPlanData[72:83,4],ImportedPlanData[72:83,5],ImportedPlanData[72:83,6],ImportedPlanData[72:83,7])
YearPlanVector <- c(ImportedPlanData[86,2],ImportedPlanData[86,3],ImportedPlanData[86,4],ImportedPlanData[86,5],ImportedPlanData[86,6],ImportedPlanData[86,7])
PlanVector_2014 <- c(ImportedPlanData[2:13,8])

# The data is saved as a vector and needs to be converted into a time series

TotalAsIs<- ts(TotalAsIsVector , start=c(2008,1), end=c(2013,12), frequency=12)
EfakAsIs <- ts(EfakAsIsVector , start=c(2008,1), end=c(2013,12), frequency=12)
WugeAsIs <- ts(WugeAsIsVector, start=c(2008,1), end=c(2013,12), frequency=12)
TotalEtelAsIs<- ts(TotalEtelAsIsVector, start=c(2008,1), end=c(2013,12), frequency=12)
BlueEtelAsIs <- ts(BlueEtelAsIsVector, start=c(2008,1), end=c(2013,12), frequency=12)
RedEtelAsIs <- ts(RedEtelAsIsVector, start=c(2008,1), end=c(2013,12), frequency=12)
YearAsIs <- ts(YearAsIsVector, start=c(2008,1), end=c(2013,12), frequency=12)
TotalAsIs_2014 <- ts(TotalAsIsVector_2014, start=c(2014,1), end=c(2014,12), frequency=12)

TotalPlan <- ts(PlanVector , start=c(2008,1), end=c(2013,12), frequency=12)
EfakPlan <- ts(EfakPlanVector, start=c(2008,1), end=c(2013,12), frequency=12)
WugePlan <- ts(WugePlanVector, start=c(2008,1), end=c(2013,12), frequency=12)
TotalEtelPlan <- ts(TotalEtelPlanVector, start=c(2008,1), end=c(2013,12), frequency=12)
BlueEtelPlan <- ts(BlueEtelPlanVector, start=c(2008,1), end=c(2013,12), frequency=12)
RedEtelPlan <- ts(RedEtelPlanVector, start=c(2008,1), end=c(2013,12), frequency=12)
YearPlan <- ts(YearPlanVector, start=c(2008,1), end=c(2013,12), frequency=12)
TotalPlan_2014 <- ts(PlanVector_2014, start=c(2014,1), end=c(2014,12), frequency=12)

# Call up the time series to check everything has worked.

TotalAsIs
EfakAsIs
WugeAsIs
TotalEtelAsIs
BlueEtelAsIs 
RedEtelAsIs 
YearAsIs
TotalAsIs_2014

TotalPlan
EfakPlan
WugePlan 
TotalEtelPlan
BlueEtelPlan
RedEtelPlan
YearPlan
TotalPlan_2014

```

##Plot Them
```{r}
par(mfrow=c(3,2))

plot(TotalAsIs, col="black", main="TotalAsIs")
plot(EfakAsIs , col="red",main="EfakAsIs")
plot(WugeAsIs, col="blue", main="WugeAsIs")
plot(TotalEtelAsIs, col="green",main="TotalEtelAsIs")
plot(BlueEtelAsIs, col="orange", main="BlueEtelAsIs")
plot(RedEtelAsIs, col="purple", main="RedEtelAsIs")
plot(TotalPlan , col="black", main="TotalPlan")
plot(EfakPlan , col="red",main="EfakPlan")
plot(WugePlan, col="blue", main="WugePlan")
plot(TotalEtelPlan, col="green",main="TotalEtelPlan")
plot(BlueEtelPlan, col="orange", main="BlueEtelPlan")
plot(RedEtelPlan, col="purple", main="RedEtelPlan")
```

# Brian Portion: Test Correlation/Model
```{r}
cor(TotalAsIs, TotalPlan )
cor(EfakAsIs , EfakPlan)
cor(WugeAsIs, WugePlan)
cor(TotalEtelAsIs, TotalEtelPlan)
cor(BlueEtelAsIs , BlueEtelPlan)
cor(RedEtelAsIs , RedEtelPlan)
cor(YearAsIs, YearPlan)
TotalAsIs_lm <- lm(TotalAsIs ~ TotalPlan , data = TotalAsIs)
summary(TotalAsIs_lm)

TotalAsIs_tslm <- tslm(TotalAsIs ~ TotalPlan )
summary(TotalAsIs_tslm)

# National Holidays
NationalHolidaysVector <- c(ImportedIndicators[170:181,2],ImportedIndicators[170:181,3],ImportedIndicators[170:181,4],ImportedIndicators[170:181,5],ImportedIndicators[170:181,6],ImportedIndicators[170:181,7])
NationalHolidays <- ts(NationalHolidaysVector, start=c(2008,1), end=c(2013,12), frequency=12)
plot(NationalHolidays, main="NationalHolidays")

cor(TotalAsIs, NationalHolidays)
cor(EfakAsIs , NationalHolidays)
cor(WugeAsIs, NationalHolidays)
cor(TotalEtelAsIs, NationalHolidays)
cor(BlueEtelAsIs , NationalHolidays)
cor(RedEtelAsIs , NationalHolidays)

# The months April and December do not correlate well with the exports data. 
# However later tests will show that these are worth considering. 
# The missing correlation is just due to the sparse structure of the NationalHolidays time series.
```

# Tony's Portion: Forecasting Models
```{r}
#################################################################################
#  7.1 Exponential Smoothing                                                    #
#################################################################################

# Exponential Smoothing uses past values to calculate a forecast. The strength 
# with which each value influences the forecast is weakened with help of a 
# smoothing parameter. Thus we are dealing with a weighted average, whose 
# values fade out the longer ago they were in the past.

#################################################################################
#  7.1.1 Simple expontential smoothing                                          #
#################################################################################

# Formula: ses(). It must be decided if alpha (the smoothing parameter
# should be automatically calculated. If initial=simple, the alpha value can 
# be set to any chosen value, if initial=optimal (or nothing, as this is the 
# default), alpha will be set to the optimal value based on ets().
# h=12 gives the number of cycles for the forecast.

Model_ses <- ses(TotalAsIs, h=12)
summary(Model_ses)
plot(Model_ses)

# The Akaike's Information Criterion(AIC/AICc) or the Bayesian Information 
# Criterion (BIC) should be at minimum.

plot(Model_ses, plot.conf=FALSE, ylab="Exports Chulwalar  )", xlab="Year", main="", fcol="white", type="o")
lines(fitted(Model_ses), col="green", type="o")
lines(Model_ses$mean, col="blue", type="o")
legend("topleft",lty=1, col=c(1,"green"), c("data", expression(alpha == 0.671)),pch=1)


#################################################################################
#  7.1.2 Holt's linear trend method                                             #
#################################################################################

# Holt added to the model in order to forecast using trends as well.
# For this it is necessary to add a beta, which determines the trend.
# If neither alpha nor beta is stated, both parameters will be optimised
# using ets(). 

Model_holt_1 <- holt(TotalAsIs,h=12)
summary(Model_holt_1)
plot(Model_holt_1)

# The trend is exponential if the intercepts(level) and the gradient (slope) are
# multiplied with eachother. The values are worse. As the Beta was very low in 
# the optimisation, the forecast is very similar to the ses() model. 


Model_holt_2<- holt(TotalAsIs, exponential=TRUE,h=12)
summary(Model_holt_2)
plot(Model_holt_2)

# As such simple trends tend to forecast the future to positively, we have added
# a dampener.
# Similar values to that of Model_holt_1 

Model_holt_3 <- holt(TotalAsIs, damped=TRUE,h=12)
summary(Model_holt_3)
plot(Model_holt_3)

# This also works for exponential trends. 
# The values remain worse. 

Model_holt_4 <- holt(TotalAsIs, exponential=TRUE, damped=TRUE,h=12)
summary(Model_holt_4)
plot(Model_holt_4)


# level and slope can be plotted individually for each model. 
plot(Model_holt_1$model$state)
plot(Model_holt_2$model$state)
plot(Model_holt_3$model$state)
plot(Model_holt_4$model$state)

plot(Model_holt_1, plot.conf=FALSE, ylab="Exports Chulwalar  )", xlab="Year", main="", fcol="white", type="o")
lines(fitted(Model_ses), col="purple", type="o")
lines(fitted(Model_holt_1), col="blue", type="o")
lines(fitted(Model_holt_2), col="red", type="o")
lines(fitted(Model_holt_3), col="green", type="o")
lines(fitted(Model_holt_4), col="orange", type="o")
lines(Model_ses$mean, col="purple", type="o")
lines(Model_holt_1$mean, col="blue", type="o")
lines(Model_holt_2$mean, col="red", type="o")
lines(Model_holt_3$mean, col="green", type="o")
lines(Model_holt_4$mean, col="orange", type="o")
legend("topleft",lty=1, col=c(1,"purple","blue","red","green","orange"), c("data", "SES","Holts auto", "Exponential", "Additive Damped", "Multiplicative Damped"),pch=1)

# As these forecasts are not very convincing at the moment, there is no need 
# to export the data.
#################################################################################
#  7.1.3 Holt-Winter's seasonal method                                          #
#################################################################################

# Holt and Winters have expanded Holt's model further to include the
# seasonality aspect. The parameter gamma, which is for smoothing the
# seasonality, was added to achieve this. The values are better than 
# the models without seasonality. This logical matches our results from the regression approaches, 
# the data is strongly influenced by seasonality. 
# In the following model, none of the parameters are given so that they
# will be optimised automatically. There are two models: one using
# an additive error model method and one using a multiplicative error model.

Model_hw_1 <- hw(TotalAsIs ,seasonal="additive",h=12)
summary(Model_hw_1)
plot(Model_hw_1)
#     AIC     AICc      BIC 
#2127.984 2137.875 2164.411 

Model_hw_2 <- hw(TotalAsIs ,seasonal="multiplicative",h=12)
summary(Model_hw_2)
plot(Model_hw_2)
#     AIC     AICc      BIC 
#2137.673 2147.564 2174.100 

# The additive model gives slightly better results than the multiplicative model.

plot(Model_hw_1, ylab="Exports Chulwalar  ", plot.conf=FALSE, type="o", fcol="white", xlab="Year")
lines(fitted(Model_hw_1), col="red", lty=2)
lines(fitted(Model_hw_2), col="green", lty=2)
lines(Model_hw_1$mean, type="o", col="red")
lines(Model_hw_2$mean, type="o", col="green")
legend("topleft",lty=1, pch=1, col=1:3, c("data","Holt Winters' Additive","Holt Winters' Multiplicative"))
```

### SES AIC:       2230.058
### SES RMSE:    609507.0
### HOLT AIC:      2235.888
### HOLT RMSE:   608787.2
### HOLT_E AIC:    2253.216
### HOLT_E RMSE: 609906.7
### HW_A AIC:     2124.856 HW Additive Model said to be best choice
### HW_A RMSE:  241685.0
### HW_M AIC:     2128.303
### HW_M RMSE:  235296.6
### Provided case says the HW additive model gives slightly better results than the multiplicative model. However, one can make a case for the multiplicative model as AICs are similar and the RMSE for the multiplicative model is 3% less than the additive.

# Araya Portion: How well does forecast meet planned exports? (May want to try this for both the HW_A and HW_M). I believe we should also try with Brian's portion as well, once he has completed it.